---
title: "Cherry Blossom Competition"
author: "Andrey"
date: "01/01/2023"
output:
  html_document:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      out.width = '80%')
```

```{r, eval=FALSE}
# Install Package
install.packages('tidyverse')
```

```{r}
# Load Package
library(tidyverse)
```


The six columns in each data file are

* _location_ a human-readable location identifier (`string`).
* _lat_ (approximate) latitude of the cherry trees (`double`).
* _long_ (approximate) longitude of the cherry trees (`double`).
* _alt_ (approximate) altitude of the cherry trees (`double`).
* _year_ year of the observation (`integer`).
* *bloom_date* date of peak bloom of the cherry trees (ISO 8601 date `string`). The "peak bloom date" may be defined differently for different sites
* *bloom_doy* days since January 1st of the year until peak bloom (`integer`). January 1st corresponds to `1`.

```{r}
cherry <- read.csv("data/washingtondc.csv") %>% 
  bind_rows(read.csv("data/liestal.csv")) %>% 
  bind_rows(read.csv("data/kyoto.csv")) %>% 
  bind_rows(read.csv("data/japan.csv")) %>%
  bind_rows(read.csv("data/meteoswiss.csv")) %>%
  bind_rows(read.csv("data/south_korea.csv")) %>%
  bind_rows(read.csv("data/vancouver.csv"))
```

```{r}
cherry
```


# Visualizing Data
```{r, fig.width=8, fig.height=3, out.width='100%', fig.cap="Time series of peak bloom of cherry trees since 1880 at three different sites."}
cherry %>% 
  filter(year >= 1990,location == "washingtondc") %>%
  ggplot(aes(x = year, y = bloom_doy)) +
  geom_point() +
  geom_step(linetype = 'dotted', color = 'gray50') +
  scale_x_continuous(breaks = seq(1990, 2020, by = 20)) +
  facet_grid(cols = vars(str_to_title(location))) +
  labs(x = "Year", y = "Peak bloom (days since Jan 1st)")
```

```{r}
cherry2 <- cherry %>% select(-c("alt","bloom_date","location"))
cherry2 <- cherry2 %>% filter(year >= 1990)
cherry2
```

```{r}
# RNOAA Package
install.packages("rnoaa")
library(rnoaa)
```
```{r}
cherry3 <- cherry2
cherry3
```


```{r}
stations = unique(cherry3[1:2])
colnames(stations) = c("latitude","longitude")
stations['id'] = 1:nrow(stations)
stations
```

```{r}
station_data <- ghcnd_stations()
```

```{r}
nearby_stations <-  meteo_nearby_stations(lat_lon_df = stations,
                    station_data = station_data, year_min=1990,year_max=2023,var=c("TMAX","TMIN"),limit=1)
```

```{r}
nearby_stations = Reduce(function(x, y) merge(x, y, all=TRUE), nearby_stations)
nearby_stations
```



```{r}
stations['nearby'] = nearby_stations[,1]
stations = stations %>% select(-c("id"))
stations
```

```{r}
# US Washington (Fix)
stations[1,"nearby"]= "USC00186350"
stations
```


```{r}
cherry4 <- left_join(x=cherry2,y=stations, by=join_by(lat==latitude,long==longitude))
```

```{r}
cherry4
```


```{r}
cherrytemp <- unique(cherry4$nearby)
cherrytemp
```


```{r, warning=FALSE}
temp4 = data.frame()
for(i in 1:length(cherrytemp)){
temp <- ghcnd_search(stationid = cherrytemp[i], var = c("tmax"), 
               date_min = "1980-01-01", date_max = "2023-02-27")[[1]] %>% mutate(year = as.integer(format(date, "%Y")),
         month = as.integer(strftime(date, '%m')) %% 12, # make December "0"
         season = cut(month, breaks = c(0, 2, 5, 8, 11),
                      include.lowest = TRUE,
                      labels = c("Winter", "Spring", "Summer", "Fall")),
         year = if_else(month == 0, year + 1L, year)) %>%
  group_by(year, season) %>%
  summarize(tmax_avg = mean(tmax, na.rm = TRUE))
temp2 <- ghcnd_search(stationid = cherrytemp[i], var = c("tmin"), 
              date_min = "1980-01-01", date_max = "2023-02-27")[[1]] %>% mutate(year = as.integer(format(date, "%Y")),
         month = as.integer(strftime(date, '%m')) %% 12, # make December "0"
         season = cut(month, breaks = c(0, 2, 5, 8, 11),
                      include.lowest = TRUE,
                      labels = c("Winter", "Spring", "Summer", "Fall")),
         year = if_else(month == 0, year + 1L, year)) %>%
  group_by(year, season) %>%
  summarize(tmin_avg = mean(tmin, na.rm = TRUE))

temp3 = merge(temp,temp2,by=c("year","season"))
temp3['stationid'] = cherrytemp[i]
temp4 = rbind(temp4,temp3)
}
```
```{r}
temp5 = temp4
temp5 = temp5[temp5$season == "Winter",]
temp5
temp5$lag10_tmax <- lag(temp5$tmax_avg, 10)
temp5$lag10_tmin <- lag(temp5$tmin_avg, 10)
temp5 = temp5 %>% select(-c("tmax_avg","tmin_avg","season"))
temp5 = temp5 %>% drop_na()
temp5
```

```{r}
cherry5 <- left_join(x=cherry4,y=temp5, by=join_by(nearby==stationid,year==year))
#cherry5 = cherry5 %>% mutate(dummy=1) %>% spread(key=season,value=dummy, fill=0)
cherry5 = cherry5 %>% drop_na()
cherry5 = cherry5 %>% select(-c("nearby"))
cherry5
```


```{r}
library(forecast)
myarima <- auto.arima(cherry5$bloom_doy,xreg=data.matrix(cherry5[,-4]))
myarima
```
```{r}
cherry5[,-4][cherry5$lat == 35.0119831 & cherry5$year==2022,]
```
```{r}
cherry_pred <- left_join(x=cherry4,y=temp4[temp4$season == "Winter",] %>% select(-c("season")), by=join_by(nearby==stationid,year==year))
cherry_pred = cherry_pred %>% filter(year > 2012)
cherry_pred
```


```{r}
#Kyoto
kyoto_x = cherry_pred[cherry_pred$lat == "35.0119831",]
kyoto_x = kyoto_x %>% select(-c("bloom_doy","nearby"))
kyoto_x$year = kyoto_x$year + 10
colnames(kyoto_x)[colnames(kyoto_x) == "tmax_avg"] ="lag10_tmax"
colnames(kyoto_x)[colnames(kyoto_x) == "tmin_avg"] ="lag10_tmin"
kyoto_x = data.matrix(kyoto_x,rownames.force=FALSE)
kyoto_x

#Liestal
liestal_x = cherry_pred[cherry_pred$lat == "47.4814",]
liestal_x = liestal_x %>% select(-c("bloom_doy","nearby"))
liestal_x$year = liestal_x$year + 10
colnames(liestal_x)[colnames(liestal_x) == "tmax_avg"] ="lag10_tmax"
colnames(liestal_x)[colnames(liestal_x) == "tmin_avg"] ="lag10_tmin"
liestal_x = data.matrix(liestal_x,rownames.force=FALSE)
liestal_x

# Washington
washingtondc_x = cherry_pred[cherry_pred$lat == "38.8853496",]
washingtondc_x = washingtondc_x %>% select(-c("bloom_doy","nearby"))
washingtondc_x$year = washingtondc_x$year + 10
colnames(washingtondc_x)[colnames(washingtondc_x) == "tmax_avg"] ="lag10_tmax"
colnames(washingtondc_x)[colnames(washingtondc_x) == "tmin_avg"] ="lag10_tmin"
washingtondc_x = data.matrix(washingtondc_x,rownames.force=FALSE)
washingtondc_x
```

```{r}
dim(washingtondc_x)
```

```{r}
#Vancouver
vancouver_x = cherry_pred[cherry_pred$lat == "49.2237",]
vancouver_x = vancouver_x %>% select(-c("bloom_doy","nearby"))
washingtondc_x$year = vancouver_x$year + 10
colnames(vancouver_x)[colnames(vancouver_x) == "tmax_avg"] ="lag10_tmax"
colnames(vancouver_x)[colnames(vancouver_x) == "tmin_avg"] ="lag10_tmin"
vancouver_x = rbind(vancouver_x,vancouver_x,vancouver_x,vancouver_x,vancouver_x,vancouver_x,vancouver_x,vancouver_x,vancouver_x,vancouver_x)
vancouver_x$year = 2023:2032
vancouver_x = data.matrix(vancouver_x,rownames.force=FALSE)
vancouver_x
```


```{r}
# Predict Kyoto
kyoto = predict(myarima,newxreg=kyoto_x)
kyotopred = round(kyoto$pred)
kyotopred

# Predict Liestal
liestal = predict(myarima,newxreg=liestal_x)
liestalpred = round(liestal$pred)
liestalpred

# Predict Washington DC
washington = predict(myarima,newxreg=washingtondc_x)
washingtonpred = round(washington$pred)
washingtonpred

# Predict Vancouver
vancouver = predict(myarima,newxreg=vancouver_x)
vancouverpred = round(vancouver$pred)
vancouverpred
```

```{r}
predictions <- data.frame(2023:2032,kyotopred,liestalpred,washingtonpred,vancouverpred)
predictions
```


```{r}
# Submission Predictions
submission_predictions = predictions
colnames(submission_predictions) = c("year","kyoto","liestal","washingtondc","vancouver")
submission_predictions
```

```{r, eval=FALSE}
write.csv(submission_predictions, file = "cherry-predictions.csv",
          row.names = FALSE)
```

